Minimax-optimal Laplacian Eigenmaps regression over Sobolev Spaces with Neighborhood Graphs

### 2021-05-26 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Updated figures for Laplacian Eigenmaps paper.

In our first experiment, we compare the mean-squared error of Laplacian eigenmaps to that of its classical counterpart. We vary the sample size from $n = 200$ to $n = 2000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\sqrt{n} \cdot \sum_{k = 1}^{n/20} k^{-s/d} \psi_k$ (Note that $f_0$ changes with $n$, but always belongs to the Sobolev ball~\textcolor{red}{(?)}.) In Figure~\ref{fig:fig1} we show the mean-squared error of Laplacian eigenmaps and spectral projection as a function of $n$, for different dimensions $d$ and order of smoothness $s$. We see that both estimators have mean-squared error converging to zero at least as fast as the optimal rate. (Here the regression function $f_0$ is a little bit smoother than the worst-case Sobolev function, which explains why the minimax rate is not always saturated). Surprisingly, we also find that Laplacian Eigenmaps slightly but consistently outperforms its continuum counterpart. 
\begin{figure*}[tb]
	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_2d.pdf}
	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_3d.pdf} 
	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_2d_2s.pdf}
	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_3d_2s.pdf} 
	\caption{Mean squared error of Laplacian eigenmaps (\texttt{LE}) and a spectral projection estimator (\texttt{SP}) as a function of sample size $n$. Each plot is on the log-log scale, and the results are averaged over 50 repetitions. Both estimators are tuned for optimal average mean squared error. The black line shows the minimax rate (in slope only; the intercept is chosen to match the observed error).}
	\label{fig:fig1}
\end{figure*}

In Figure~\ref{fig:fig2} we compare tests using Laplacian eigenmaps and spectral projection. \textcolor{red}{(TODO): Need to ask Siva for help.}
\begin{figure*}[tb]
	\caption{Worst-case risk Laplacian eigenmaps (\texttt{LE}) and spectral projection (\texttt{SP}) tests, as a function of sample size $n$.}
	\label{fig:fig2}
\end{figure*}

@@

\begin{figure*}[tb]
	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters/mse_by_number_of_eigenvectors.pdf}
	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters/mse_by_radius.pdf} 
	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters_2s/mse_by_number_of_eigenvectors.pdf}
	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters_2s/mse_by_radius.pdf} 
	\caption{Mean squared error of Laplacian Eigenmaps as a function of hyperparameters. For all experiments, sample size $n = 2000$, and the results are averaged over $50$ repetitions. Colors show~\textcolor{red}{(?)}}
	\label{fig:fig3}
\end{figure*}

>>>>

In our first experiment, we compare the mean-squared error of Laplacian eigenmaps to that of its classical counterpart. We vary the sample size from $n = 200$ to $n = 2000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\sqrt{n} \cdot \sum_{k = 1}^{n/20} k^{-s/d} \psi_k$ (Note that $f_0$ changes with $n$, but always belongs to the Sobolev ball~\textcolor{red}{(?)}.) In Figure~\ref{fig:fig1} we show the mean-squared error of Laplacian eigenmaps and spectral projection as a function of $n$, for different dimensions $d$ and order of smoothness $s$. We see that both estimators have mean-squared error converging to zero at least as fast as the optimal rate. (Here the regression function $f_0$ is a little bit smoother than the worst-case Sobolev function, which explains why the minimax rate is not always saturated). Surprisingly, we also find that Laplacian Eigenmaps slightly but consistently outperforms its continuum counterpart. 
\begin{figure*}[tb]
	\includegraphics[width=.245\textwidth]{figures/in_sample_mse/mse_by_sample_size_1d_1s.pdf}
	\includegraphics[width=.245\textwidth]{figures/in_sample_mse/mse_by_sample_size_1d_2s.pdf}
	\includegraphics[width=.245\textwidth]{figures/in_sample_mse/mse_by_sample_size_2d_1s.pdf}
	\includegraphics[width=.245\textwidth]{figures/in_sample_mse/mse_by_sample_size_2d_2s.pdf}
	\caption{Mean squared error of Laplacian eigenmaps (\texttt{LE}) and a spectral projection estimator (\texttt{SP}) as a function of sample size $n$. Each plot is on the log-log scale, and the results are averaged over 400 repetitions. Both estimators are tuned for optimal average mean squared error. The black line shows the minimax rate (in slope only; the intercept is chosen to match the observed error).}
	\label{fig:fig1}
\end{figure*}

We also compare the error, over a held-out test set, of Laplacian eigenmaps plus kernel smoothing to the spectral projection estimator. The relative performances of the two estimators, as seen in Figure~\ref{fig:fig2}, is very similar to that seen in Figure~\ref{fig:fig1}, supporting our theoretical claim that the additional error incurred by kernel smoothing of Laplacian Eigenmaps is negligible.
\begin{figure*}[tb]
	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_1d_1s.pdf}
	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_1d_2s.pdf}
	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_2d_1s.pdf}
	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_2d_2s.pdf}
	\caption{Mean squared error of Laplacian eigenmaps plus kernel smoothing (\texttt{LE+KS}) and a spectral projection estimator (\texttt{SP}) as a function of sample size $n$. Each plot is on the log-log scale, and the results are averaged over 200 repetitions. Both estimators are tuned for optimal average mean squared error. The black line shows the minimax rate (in slope only; the intercept is chosen to match the observed error).}
	\label{fig:fig2}
\end{figure*}

In Figure~\ref{fig:fig3} we compare tests using Laplacian eigenmaps and spectral projection. \textcolor{red}{(TODO): Need to ask Siva for help.}
\begin{figure*}[tb]
	\caption{Worst-case risk Laplacian eigenmaps (\texttt{LE}) and spectral projection (\texttt{SP}) tests, as a function of sample size $n$.}
	\label{fig:fig3}
\end{figure*}

@@

\begin{figure*}[tb]
	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters/mse_by_number_of_eigenvectors.pdf}
	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters/mse_by_radius.pdf} 
	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters_2s/mse_by_number_of_eigenvectors.pdf}
	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters_2s/mse_by_radius.pdf} 
	\caption{Mean squared error of Laplacian Eigenmaps as a function of hyperparameters. For all experiments, sample size $n = 2000$, and the results are averaged over $200$ repetitions. Colors show~\textcolor{red}{(?)}}
	\label{fig:fig4}
\end{figure*}
